---
title: "GMLA"
author: "Matthew Shisler, Kate Sanborn, Richard Watson"
date: "11/25/2021"
output: word_document
---

In conjunction with the tutorial version, we created this more automated way of 
running through the Gradient Maximum Likelihood Algorithm.

The end result is a graph with pasted scores, the true source, the estimated source, 
and the distance between the true source and estimated source.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(tidygraph)
library(ggraph)
library(netdiffuseR)
library(EpiModel)
library(tidyverse)
library(shp2graph)
```
This utilizes the Duke propagation tutorial code referenced in the GMLA tutorial.
```{r}
n.people <- 100
p.infection <- 0.5
contact.rate <- 0.05  # prob. of contact between any 2 people in the network

set.seed(919)

# generate the network
net <- play_erdos_renyi(n.people,contact.rate, directed = F)
ggraph(net, layout = "kk") +
  geom_edge_link() +
  geom_node_point(size = 6)

# choose ONE infected person
infected <- rep(FALSE, n.people)
infected[sample(1:n.people,1)] <- TRUE

# extract the edge list for the sim
el <- net %>% activate(edges) %>% as.data.frame()

# pre-allocate a list of the sim. Each element is a logical
# vector for nodes being infected. The list is indexed by time.
# right now the max is set to 1000. The list will grow on its
# own if full infection takes longer.
infections <- vector(length = 1000, mode = "list")
infections[[1]] <- infected

# loop infection steps until EVERYONE is infected
# (rather than loop for a pre-sprecified time)
t=2
while (sum(!infections[[t-1]])>0) {
  infections[[t]] <- infections[[t - 1]]
  
  # Pick an edge at random
  random.edge <- sample(nrow(el), size = 1)
  
  # If the edge is discordant, flip a coin to decide if the infection spreads
  if (with(el[random.edge, ], 
           infections[[t]][from] != infections[[t]][to])) {
    
    who.susceptible <- with(
      el[random.edge, ],
      c(from, to)[!c(infections[[t]][from], infections[[t]][to])]
      )
    
    infections[[t]][who.susceptible] <- sample(
      c(T, F), 
      size = 1, 
      prob = c(p.infection, 1 - p.infection)
    )
  }
  t=t+1
}

# Record the time it took for everyone to become infected
max.time=t-1

# trim the infections list of any extra empy list elements
# (because we pre-allocated too many)
infections <- infections[lapply(infections,length)>0]

# transform the infections list into a dataframe which
# stores the nodes ids and the time that they were infected
# with respect to the BEGINNING of the simulation.
results <- infections %>%
  lapply(FUN = as.data.frame) %>%
  lapply(FUN = magrittr::set_names, value = "infected") %>%
  bind_rows(.id = "t") %>%
  mutate(id = rep(1:n.people, times = max.time),
         t = as.integer(t)) %>%
  tbl_df %>%   group_by(id) %>%
  summarize(time.infected = min(t[infected]))

# Add the infections times and the source identifier to
# the network object
net <- net %>%
  activate(nodes) %>%
  mutate(node_id = 1:n.people) %>%
  mutate(true_infection_time = results$time.infected) %>%
  mutate(source = infected)

# graph the network with a filter on infection time.
# it'd be kind of neat if we could animate this. . .
ggraph(net, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=true_infection_time<100), size = 6)

```
Next task is to try to convert Kate's code so far into something that works with the tidygraph structure.
Hopefully this will be in general easier to operate on.
```{r}
# identify proportion of nodes to be observers
prop_observers <- 0.2
observers_index <- sample(1:n.people, 
                          ceiling(n.people*prop_observers), 
                          replace=F)

# create new columns identifying observers and the relative infection time among observers.
net <- net %>%
  activate(nodes) %>%
  mutate(observer = ifelse(node_id %in% observers_index,T,F)) %>%
  mutate(observed_infection_time = ifelse(observer==T, 
                                          true_infection_time, 
                                          NA)) %>%
  mutate(observed_infection_time = observed_infection_time - min(observed_infection_time, 
                                                                 na.rm = T)) %>%
  arrange(desc(observer),true_infection_time)

# current structure of nodes
net %>%
  as.data.frame()
```
```{r}
# mean propagation time per edge, mu
# variance of propagation time per edge, sigma^2
# propagation ratio, lambda
mu = 1/p.infection
sigma2 <- (1-p.infection)/p.infection^2 
lambda <- mu/sqrt(sigma2)


# how many observers do we care about?
k0 <- 10
top_observers <- net %>% 
  filter(observer==T) %>% 
  pull(node_id) %>%
  head(n=k0)

d_vec <- net %>% 
  filter(observer==T) %>% 
  pull(observed_infection_time) %>%
  head(n=k0) %>% tail(n=-1)

# set the reference observer, o_K
# and other top observers, o_k (k=1,...,K-1)
reference_observer <- top_observers[1]
other_observers    <- top_observers[-1]

# we need to create the covariance matrix Lambda
# first, compute the shortest paths from the reference observer to the other top observers
lam1 <- shortest_paths(net, reference_observer, other_observers, output = "epath")$epath

# next, the diagonal of the matrix will be the size of each path multiplied by sigma^2
# this is because the intersection between a set and itself is the original set.
Lambda_mat <- diag(sigma2*as.integer(lapply(lam1, length)))


# we should only need to compute the upper triangle, right?
for (i in 1:(length(other_observers)-1)){
  for (j in (1+i):length(other_observers)){
    Lambda_mat[i,j] <- sigma2*length(intersection(lam1[[i]], lam1[[j]]))
  }
}

# make it symmetric and then take the inverse
Lambda_mat <- Lambda_mat*upper.tri(Lambda_mat, diag = F) + t(Lambda_mat)
inv_Lambda_mat <- solve(Lambda_mat)
```






```{r}
# now for the loop
v<-vector()
v[1] <- reference_observer
score_max <- 0
S_nodes  <- c() 
S_scores <- c()

# get the initial v node neighbors
neighbors <- as.integer(neighbors(net, v[1]))


k=1
while (!is_empty(neighbors)){
  # initialize neighbor scores
  neighbor_scores <- rep(0,length(neighbors))
  
  # compute the score for each neighbor
  for (i in  1:length(neighbors)){
    # get the shortest path from neighbor i to reference observers
    # short_path_to_ref <- shortest_paths(net, node, reference_observer, output = "epath")
    # short_paths_to_others <- shortest_paths(net, node, other_observers, output = "both")
    
    # we only need distances to compute mu, this function might be faster
    dist_to_ref <- as.integer(distances(net, neighbors[i], reference_observer))
    dist_to_others <- as.integer(distances(net, neighbors[i], other_observers))
    
    mu_vec <- mu*(dist_to_others-dist_to_ref)
    # do we really need to compute the denominator of the score?
    #score <- exp(-0.5*(d_vec-mu_vec)%*%inv_Lambda_mat%*%(d_vec-mu_vec))
    neighbor_scores[i] <- (-0.5*(d_vec-mu_vec)%*%inv_Lambda_mat%*%(d_vec-mu_vec)) - log(sqrt(det(Lambda_mat)))
    # get shortest paths from neighbor i to other observers
  }

  
  # determine the next v, store the nodes and scores, construct the next
  # neighborhood remembering to not consider nodes we already have a score for.
  # if this neighborhood is empty, then we can stop because all the surrounding
  # nodes have a smaller score.
  v[k+1] <- neighbors[which(neighbor_scores==max(neighbor_scores))]
  
  # don't really need the max score any more. The stopping condition is
  # if the next set of neighbors is empty.
  score_max <- max(neighbor_scores)
  
  # add nodes and score to running list.
  S_nodes <- c(S_nodes, neighbors)
  S_scores <- c(S_scores, neighbor_scores)
  
  # get the next set of neighbors less the nodes that already have scores
  neighbors <- setdiff(as.integer(neighbors(net, v[k+1])), S_nodes)
  k = k+1
}

# add scores to the network object
net <- net %>% 
  activate(nodes) %>%
  left_join(tibble(node_id = S_nodes, score = S_scores), by="node_id")

net %>%
  as.data.frame()

```

```{r}
# the node with the maximum number is the estimated source.
# in 
# graph isn't really helpful. . .messy
ggraph(net, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(colour=true_infection_time<16, shape=observer),  size = 6) +
  geom_node_text(aes(label=node_id))+
  geom_node_label(aes(label=score), nudge_y = 0.3)

# the path of the estimator (in the direction of the gradient)
# is in v.
v

# The true source and estimated source
which(infected)
v[length(v)]


# distance from true source to estimated source.
distances(net,which(infected),v[length(v)])


```


















