---
title: "Untitled"
author: "Matthew Shisler"
date: "12/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(tidygraph)
library(ggraph)
library(EpiModel)
library(tidyverse)
library(shp2graph)
library(scales)
```
The re-factored code from Duke. The issue we want to explore is in the loop that propagate the virus. This first chunk is set-up.
```{r}
n.people <- 121
p.infection <- 0.5
contact.rate <- 0.05  # prob. of contact between any 2 people in the network

set.seed(919)

# generate the network
net_er <- play_erdos_renyi(n.people,contact.rate, directed = F)
n_level <- 5   # number of levels in the tree
d       <- 3   # degree of each node in the tree (other than root)

n_nodes <- (d^n_level-1)/(d-1)  # the number of nodes for a "full" tree.

# will use tidygraph as the API for igraph.
net_tree <- create_tree(n_nodes, d, directed = F)


ggraph(net_er, layout = "kk") +
  geom_edge_link() +
  geom_node_point(size = 6)

# choose ONE infected person
infected <- rep(FALSE, n.people)
infected[sample(1:n.people,1)] <- TRUE
source_id <- which(infected==T)


# # extract the edge list for the sim
# el <- net %>% activate(edges) %>% as.data.frame()
# 
# # pre-allocate a list of the sim. Each element is a logical
# # vector for nodes being infected. The list is indexed by time.
# # right now the max is set to 1000. The list will grow on its
# # own if full infection takes longer.
# infections <- vector(length = 1000, mode = "list")
# infections[[1]] <- infected
```


This next chunk is the actual simulation. Let's describe the actions taken in a single time step. First, a random edge is drawn from the list of edges. If the edge is discordant (i.e. connecting an infected and non-infected node), then a coin is flipped to determine if the infection will propagate to the non-infected individual. If the randomly infected edge is not discordant, then we take another time step. This results in long propagation times since the action of the propagation along an edge is conditional on whether the edge was selected in the first place.

We believe that the proper actions to take in a single time step would be to flip a coin for EVERY discordant edge. This way the infection would propagate at a much more reasonable rate.

```{r}
# build graph with variables for infected nodes and discordant edges
net_sim <- net_tree %>%
  activate(nodes) %>%
  mutate(node_id = 1:n.people) %>%
  mutate(infected=infected) %>%
  mutate(time_infected = ifelse(infected,0,NA)) %>%
  activate(edges) %>%
  mutate(from_infected = infected[from]) %>%
  mutate(to_infected = infected[to]) %>%
  mutate(discordant = from_infected != to_infected)

# inspect the node and edge dataframes if desired
net_sim %>% activate(nodes) %>% as_tibble()
net_sim %>% activate(edges) %>% as_tibble()

# initialize stopping condition (all nodes infected)
infect_check <- net_sim %>% 
    activate(nodes) %>%
    pull(infected)

# initialize time step and max time steps
time_limit <- 200
time_step = 1

# propagate infection until all are infected or we reach the time limit
while(any(!infect_check) && time_step < time_limit){  
  
  # identify susceptible nodes
  sus_nodes <- net_sim %>% activate(edges) %>%
    filter(discordant) %>%
    mutate(susceptible = from*to_infected+to*from_infected) %>%
    pull(susceptible)
    
  # flip coin for all susceptible nodes
  # nodes which are susceptible to infection from more than one neighbor get separate coin flips
  # for each neighbor
  new_infected <- sus_nodes[sample(c(T,F), length(sus_nodes), replace = T, prob=c(p.infection,1-p.infection))]
  
  # update graph object with newly infected individuals
  net_sim <- net_sim %>% 
    activate(nodes) %>%
    mutate(infected = ifelse(node_id %in% new_infected,TRUE,infected)) %>%
    mutate(time_infected = ifelse(node_id %in% new_infected,time_step,time_infected)) %>%
    activate(edges) %>%
    mutate(from_infected = .N()$infected[from]) %>%
    mutate(to_infected = .N()$infected[to]) %>%
    mutate(discordant = from_infected != to_infected)
  
  # update stopping criteria and time step
  infect_check <- net_sim %>% 
    activate(nodes) %>%
    pull(infected)
  time_step = time_step + 1
}

# inspect result graph nodes and edges
net_sim %>% activate(nodes) %>% as_tibble()
net_sim %>% activate(edges) %>% as_tibble()
```

Graphing
```{r}
max_time <- max(net_sim %>% activate(nodes) %>% pull(time_infected))

# step through time
for (i in seq(from=1, max_time, length.out=15)){
gr <- ggraph(net_sim, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=time_infected<i), size = 6) +
  theme_void()
print(gr)
}

# use color gradient for single graph
ggraph(net_sim, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=time_infected), size = 6) +
  scale_color_gradient(low = "red", high = "yellow", na.value = NA) +
  labs(color="Time to infection") +
  theme_void()

```

Continuous time step approach. Much more simple. . .
```{r}
net_det <- net_tree %>%
  activate(nodes) %>%
  mutate(node_id = 1:n.people) %>%
  mutate(infected=infected) %>%
  activate(edges) %>%
  mutate(transmission_time = rexp(nrow(.E()))) %>%
  activate(nodes) %>%
  mutate(time_infected = as.numeric(distances(.,v=source_id, to=V(.), weights = E(.)$transmission_time))) %>%
  mutate(bfs_prop_time = as.numeric(distances(.,v=source_id, to=V(.), weights = NA)))

net_det %>% activate(nodes) %>% as_tibble()
net_det %>% activate(edges) %>% as_tibble()

# time step through infection
max_time <- max(net_det %>% activate(nodes) %>% pull(time_infected))
for (i in seq(from=1, max_time, length.out=15)){
gr <- ggraph(net_det, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=time_infected<i), size = 6) +
  theme_void()
print(gr)
}

# compare discrete to continuous infection simulation
ggraph(net_sim, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=rescale(time_infected)), size = 6) +
  scale_color_gradient(low = "red", high = "yellow", na.value = NA) +
  labs(color="Time to infection") +
  theme_void()

ggraph(net_det, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=rescale(time_infected)), size = 6) +
  scale_color_gradient(low = "red", high = "yellow", na.value = NA) +
  labs(color="Time to infection") +
  theme_void()
```

```{r}

net_sim <- net_sim %>%
  activate(nodes) %>%
  mutate(scaled_time_infected = rescale(time_infected))

net_det <- net_det %>%
  activate(nodes) %>%
  mutate(scaled_time_infected = rescale(time_infected)) %>%
  mutate(scaled_bfs_prop_time = rescale(bfs_prop_time))

ggraph(net_sim, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=scaled_time_infected), size = 6) +
  scale_color_gradient(low = "red", high = "yellow", na.value = NA) +
  labs(color="Time to infection") +
  theme_void()

ggraph(net_det, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=scaled_time_infected), size = 6) +
  scale_color_gradient(low = "red", high = "yellow", na.value = NA) +
  labs(color="Time to infection") +
  theme_void()

ggraph(net_det, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=scaled_bfs_prop_time), size = 6) +
  scale_color_gradient(low = "red", high = "yellow", na.value = NA) +
  labs(color="Time to infection") +
  theme_void()

```


```{r}
set.seed(2021)
demo_er <- play_erdos_renyi(20, 0.15, directed = F) %>%
  mutate(node_id = 1:20)

ggraph(demo_er, layout = "kk") +
  geom_edge_link() +
  geom_node_point(size = 8, color="cyan3") +
  geom_node_text(aes(label=node_id)) +
  theme_void()

# the shortest path from 1 to 14 requires 3 hops
shortest_paths(demo_er,1,6)

# histogram of path lengths for all simple paths from node 1 to 14.
histogram(sapply(all_simple_paths(demo_er, 1, 6),length)-1, type="count",
          xlab="Lengths (hops) of Simple Paths from Node 1 to Node 6")

path_data <- as.data.frame(table(sapply(all_simple_paths(demo_er, 1, 6),length)-1))

path_data$Var1 <- as.numeric(levels(path_data$Var1))

sampled_transmission_times <- data.frame(length=rep(path_data[,1],times=path_data[,2]), time=0)
path_length_of_shortest_time <- rep(0,1000)
transmission_time <- rep(0,1000)

for (j in 1:1000){
  for (i in 1:nrow(sampled_transmission_times)){
    sampled_transmission_times[i,2] <- rgamma(1,sampled_transmission_times[i,1])
  }
  path_length_of_shortest_time[j] <- sampled_transmission_times[which(sampled_transmission_times[,2] == min(sampled_transmission_times[,2])),1]
  transmission_time[j] <- min(sampled_transmission_times[,2])
}
  
histogram(path_length_of_shortest_time,
          breaks=seq(3,12,1),
          xlab = "Path length (hops) of minimum infection time")
histogram(transmission_time,
          xlab="Node 6 infection time")


```
Whole infection simulation
```{r}

sim_bfs_path_diff <- vector(mode="list", length=1000)

for (i in 1:1000){
# generate a random ER graph with 100 nodes and contact rate = 0.05
  net_er <- play_erdos_renyi(100,0.1, directed = F)
  
  # choose a random initial infected node
  infected <- rep(FALSE, 100)
  infected[sample(1:100,1)] <- TRUE
  source_id <- which(infected==T)
  
  # simulate the edgewise infection times and shortest weighted paths for 
  # nodewise infection times.
  net_det <- net_er %>%
    activate(nodes) %>%
    mutate(node_id = 1:100) %>%
    mutate(infected=infected) %>%
    activate(edges) %>%
    mutate(transmission_time = rexp(nrow(.E()))) %>%
    activate(nodes) %>%
    mutate(time_infected = as.numeric(distances(.,v=source_id, to=V(.), weights = E(.)$transmission_time))) %>%
    mutate(bfs_prop_time = as.numeric(distances(.,v=source_id, to=V(.), weights = NA)))
  
  # compute the number of hops in the BFS paths and the paths resulting from the simulation.
  bfs_path_lengths<- sapply(shortest_paths(net_det, from=source_id,to=V(net_det), weights = NA, output = "epath")$epath, length)
  sim_path_lengths<- sapply(shortest_paths(net_det, from=source_id,to=V(net_det), weights = E(net_det)$transmission_time, output = "epath")$epath, length)
  
  # histogram(bfs_path_lengths,
  #           breaks = seq(0,max(sim_path_lengths),1),
  #           ylim=c(0,50))
  # histogram(sim_path_lengths,
  #           breaks = seq(0,max(sim_path_lengths),1),
  #           ylim=c(0,50))
  # histogram(sim_path_lengths-bfs_path_lengths,
  #           breaks=seq(0,7,1))
  
  sim_bfs_path_diff[[i]] <- unclass(table(sim_path_lengths-bfs_path_lengths))
  
}  


results <- plyr::ldply(sim_bfs_path_diff,rbind)



histogram(colSums(results, na.rm = T))

ggplot(colSums(results, na.rm = T)) + geom_bar(aes(x=1:12,y=))

data <- data.frame(x=0:13, counts=colSums(results, na.rm = T))
data[1,2] <- data[1,2]-1000

ggplot(data) + geom_bar(aes(x=x,y=counts), stat="identity") +
  xlab("Difference in BFS path hops and actual path hops in 1000 SI sims") +
  theme_bw() + theme(text = element_text(size = 15)) 

max(results[,9], na.rm = T)





```

Distribution of infection time
```{r}
net_er <- play_erdos_renyi(100,0.05, directed = F)

  infected <- rep(FALSE, 100)
  infected[sample(1:100,1)] <- TRUE
  source_id <- which(infected==T)

infection_times <- vector(mode="list", length=1000)

for (i in 1:1000){
  
  # simulate the edgewise infection times and shortest weighted paths for 
  # nodewise infection times.
  time_infected <- net_er %>%
    activate(nodes) %>%
    mutate(node_id = 1:100) %>%
    mutate(infected=infected) %>%
    activate(edges) %>%
    mutate(transmission_time = rexp(nrow(.E()))) %>%
    activate(nodes) %>%
    mutate(time_infected = as.numeric(distances(.,v=source_id, to=V(.), weights = E(.)$transmission_time))) %>%
    pull(time_infected)

infection_times[[i]] <- time_infected
}
  
results <- plyr::ldply(infection_times,rbind)
results <- results %>% gather()
```
Histograms of for simulated infection time distributions of each node.
```{r}
ggplot(results[1:12000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[12001:24000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[24001:36000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[36001:48000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[48001:60000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[60001:72000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[72001:84000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[84001:96000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')

ggplot(results[96001:100000,], aes(value)) +
  geom_histogram() +
  facet_wrap(~key, scales='free')
```
Many of these histograms look normal at first glance. The ones that are more highly skewed are closer to the source node. 
```{r}
# looking at larger and larger neighborhoods
ego(net_er, order=1, nodes=source_id)
ego(net_er, order=2, nodes=source_id)
ego(net_er, order=5, nodes=source_id)
ego(net_er, order=3, nodes=source_id)
ego(net_er, order=4, nodes=source_id)
ego(net_er, order=5, nodes=source_id)

```

As k increases, the gamma distribution approaches the normal distribution.
```{r}

k = 25

x1 <- seq(0,50, length.out=1000)
y11 <- dgamma(x1, k, rate=1)
y12 <- dnorm(x1, k,sqrt(k))

plot(x1, y11, type="l")
lines(x1, y12, col=2)
```

Writing the discrete sim as a function - easily handles multiple sources, no delay times for edges must be calculated from output. Scaled on arbitrary time steps.
```{r}

sim_SI_discrete <- function(net,source_id,infection_prob=0.5,time_limit=200){
  # initialize
  net_sim <- net %>%
    activate(nodes) %>%
    mutate(node_id = 1:(nrow(.N()))) %>%
    mutate(infected=FALSE) %>%
    mutate(infected=ifelse(node_id %in% source_id,TRUE,infected)) %>%
    mutate(time_infected = ifelse(infected,0,NA)) %>%
    activate(edges) %>%
    mutate(from_infected = .N()$infected[from]) %>%
    mutate(to_infected = .N()$infected[to]) %>%
    mutate(discordant = from_infected != to_infected)
  
  # initialize stopping condition (all nodes infected)
  infect_check <- net_sim %>% 
      activate(nodes) %>%
      pull(infected)
  
  # initialize time step
  time_step = 1
  # propagate infection until all are infected or we reach the time limit
  while(any(!infect_check) && time_step < time_limit){  
    
    # identify susceptible nodes
    sus_nodes <- net_sim %>% activate(edges) %>%
      filter(discordant) %>%
      mutate(susceptible = from*to_infected+to*from_infected) %>%
      pull(susceptible)
      
    # flip coin for all susceptible nodes
    # nodes which are susceptible to infection from more than one neighbor get separate coin flips
    # for each neighbor
    new_infected <- sus_nodes[sample(c(T,F), 
                                     length(sus_nodes), 
                                     replace = T, 
                                     prob=c(infection_prob,1-infection_prob))]
    
    # update graph object with newly infected individuals
    net_sim <- net_sim %>% 
      activate(nodes) %>%
      mutate(infected = ifelse(node_id %in% new_infected,TRUE,infected)) %>%
      mutate(time_infected = ifelse(node_id %in% new_infected,time_step,time_infected)) %>%
      activate(edges) %>%
      mutate(from_infected = .N()$infected[from]) %>%
      mutate(to_infected = .N()$infected[to]) %>%
      mutate(discordant = from_infected != to_infected)
    
    # update stopping criteria and time step
    infect_check <- net_sim %>% 
      activate(nodes) %>%
      pull(infected)
    time_step = time_step + 1
  }
  
  net_sim <- net_sim %>%
    activate(edges) %>%
    dplyr::select(-c(from_infected, to_infected, discordant))
  
  return(list(net_sim = net_sim, source_id = source_id, infection_prob = infection_prob, time_steps = time_step))
}

out <- sim_SI_discrete(net_er, 14, infection_prob = 0.05)

out$net_sim %>% activate(nodes) %>% as_tibble()
out$net_sim %>% activate(edges) %>% as_tibble()

max_time <- max(out$net_sim %>% activate(nodes) %>% pull(time_infected))+1
# step through time
for (i in seq(from=1, max_time, length.out=15)){
gr <- ggraph(out$net_sim, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=time_infected<i), size = 6) +
  theme_void()
print(gr)
}
```

Continuous sim as a function - no discrete time steps. MUCH MUCH faster. currently only supports single source. Infection times readily interpreted.
```{r}
sim_sI_continuous <- function(net, source_id, infection_rate=1){
  
  net_sim <- net %>%
    activate(nodes) %>%
    mutate(node_id = 1:(nrow(.N()))) %>%
    mutate(infected=FALSE) %>%
    mutate(infected=ifelse(node_id %in% source_id,TRUE,infected)) %>%
    activate(edges) %>%
    mutate(transmission_time = rexp(nrow(.E()), rate = infection_rate)) %>%
    activate(nodes) %>%
    mutate(time_infected = as.numeric(distances(.,v=source_id, to=V(.), weights = E(.)$transmission_time)))%>%
    mutate(infected = ifelse(time_infected>=0,TRUE,FALSE))
  
  
  return(list(net_sim = net_sim, source_id = source_id, infection_rate = infection_rate))
}

out <- sim_sI_continuous(net_er, c(14), infection_rate= 1)

out$net_sim %>% activate(nodes) %>% as_tibble()
out$net_sim %>% activate(edges) %>% as_tibble()

max_time <- max(out$net_sim %>% activate(nodes) %>% pull(time_infected))+1
# step through time
for (i in seq(from=1, max_time, length.out=15)){
gr <- ggraph(out$net_sim, layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(color=time_infected<i), size = 6) +
  theme_void()
print(gr)
}
```
Comparing computational cost of discrete and continuous sim.
```{r}
library(microbenchmark)

results <- microbenchmark(
  `discrete` = sim_SI_discrete(net_er,14),
  `continuous` = sim_sI_continuous(net_er,14),
  times=100
)

results
plot(results)

```



